A new method cluster Detection from three-dimensional weak lensing mass map

### 2021-01-13 doc/paper_ms_method_HSCY1/massmap_sim.tex
small revision on paper
some minor corrections

We sample halos in a two-dimensional redshift-mass plane. The redshift-mass
plane is evenly divided into eight redshift bins and eight mass bins. We
randomly shift the input halo redshifts and halo masses from the bins' centers
by a small amount. In the simulation, we set the non-linear over-density
($\Delta_{\rm{vir}}$) to $200$, and $M_{200}$ refers to the virial mass. The
concentration of the NFW halo is set to as a function of the halo's virial mass
($M_{200}$) and redshift ($z_{h}$) according to
\citet{c-M_Magneticum-Ragagnin2019}
\begin{equation}
c_{h}=6.02\times\left(\frac{M_{200}}{10^{13} M_{\odot}}\right)^{-0.12}
\left(\frac{1.47}{1.+z_h}\right)^{0.16}.
\end{equation}
The halos are truncated at the virial radius.
The weak-lensing shear fields of these NFW halos are simulated according to
\citet{haloModel-TJ2003-3pt}. The shear distortions are applied to one hundred
realizations of galaxy catalogs with the HSC-like shape measurement error and
photo-$z$ uncertainty.

@@

The mock galaxy catalogs are generated using the HSC S16A shape catalog
\citep{HSC1-catalog}.  We use the galaxies in a one square degree region at the
center of tract 9347 \citep{HSC1-data}.  The galaxies' positions are randomized
to distribute homogeneously  in the one-square degree stamp statistically. We
randomly assign its redshift for each galaxy following the MLZ photo-$z$
probability distribution function \citep{HSC1-photoz}.

By randomly rotating the galaxies in the shape catalog, we simulate the
HSC-like shape estimation errors with different realizations.  The histogram of
the first component of the HSC-like shape estimation error on galaxy level is
shown in Figure \ref{fig_noiseHistogram}.  The corresponding histogram of the
shape measurement error on the pixel level after the smoothing and pixelization is
also shown in \ref{fig_noiseHistogram}. The standard deviation map of the
noise is demonstrated in Figure \ref{fig_noistdmap}. As demonstrated in Figure
\ref{fig_noiseHistogram}, even though the shape measurement error on the galaxy
level does not fully follow Gaussian distribution, the error is well described
by Gaussian distribution after the smoothing and pixelization.

>>>>

We sample halos in a two-dimensional redshift-mass plane. The redshift-mass
plane is evenly divided into eight redshift bins and eight mass bins. We
randomly shift the input halo redshifts and halo masses from the bin centers
by a small amount. The concentration of the NFW halo is set to as a function of
the halo mass ($M_{200}$) and redshift ($z_{h}$) according to
\citet{c-M_Magneticum-Ragagnin2019}
\begin{equation}
c_{h}=6.02\times\left(\frac{M_{200}}{10^{13} M_{\odot}}\right)^{-0.12}
\left(\frac{1.47}{1.+z_h}\right)^{0.16}.
\end{equation}
The halos are truncated at the virial radius.
The weak-lensing shear fields of these NFW halos are simulated according to
\citet{haloModel-TJ2003-3pt}. The shear distortions are applied to one hundred
realizations of galaxy catalogs with the HSC-like shape measurement error and
photo-$z$ uncertainty.

@@

The mock galaxy catalogs are generated using the HSC S16A shape catalog
\citep{HSC1-catalog}.  We use the galaxies in a one square degree region at the
center of tract 9347 \citep{HSC1-data}. The average galaxy number density in
this region is $22.94~\rm{arcmin}^{-2}$. The positions of galaxies are randomized to
distribute homogeneously  in the one-square degree stamp statistically. We
randomly assign its redshift for each galaxy following the MLZ photo-$z$
probability distribution function \citep{HSC1-photoz}.

By randomly rotating the galaxies in the shape catalog, we simulate the
HSC-like shape estimation errors with different realizations.  The histogram of
the first component of the HSC-like shape estimation error on galaxy level is
shown in Figure \ref{fig_noiseHistogram}.  The corresponding histogram of the
shape measurement error on the pixel level after the smoothing and pixelization is
also shown in Figure \ref{fig_noiseHistogram}. The standard deviation map of the
noise is demonstrated in Figure \ref{fig_noistdmap}. As demonstrated in Figure
\ref{fig_noiseHistogram}, even though the shape measurement error on the galaxy
level does not fully follow Gaussian distribution, the error is well described
by Gaussian distribution after the smoothing and pixelization.

### 2021-01-13 doc/paper_ms_method_HSCY1/ms.tex
some minor corrections

\begin{document}
\title{Cluster Detection from three-dimensional Lensing Mass Map}
% \author{Xiangchong Li}
% \affiliation{Department of Physics, University of Tokyo, Tokyo 113-0033, Japan}
% \affiliation{Kavli Institute for the Physics and Mathematics of the Universe (WPI),\\
% University of Tokyo, Kashiwa 277-8583, Japan}
% \author{Masamune Oguri}
% \affiliation{Department of Physics, University of Tokyo, Tokyo 113-0033, Japan}
% \affiliation{Kavli Institute for the Physics and Mathematics of the Universe
% (WPI),\\
% University of Tokyo, Kashiwa 277-8583, Japan}
% \affiliation{Research Center for the Early Universe, University of Tokyo, Tokyo
% 113-0033, Japan}
% \author{Wentao Luo}
% \affiliation{Kavli Institute for the Physics and Mathematics of the Universe
% (WPI),\\
% University of Tokyo, Kashiwa 277-8583, Japan}
% \email{xiangchong.li@ipmu.jp}

\begin{abstract}
A new method is developed to reconstruct high-resoluion three-dimensional mass
maps from photometric weak-lensing shear measurements. The three-dimensional
mass map is modeled as a summation of the NFW basis atoms, which have
two-dimensional multi-scale NFW surface density profiles on the transverse
plane and one-dimensional Dirac delta functions in the line of sight direction.
The adaptive lasso algorithm is applied to find a sparse reconstruction of the
mass maps.
We study the performance of three-dimensional cluster detection from the
reconstructed mass map with simulations that applies shear distortions from
isolated halo to HSC-like galaxy shapes with realistic photometric redshift
uncertainties .
Our findings are summarized as follows:
1) The lasso reconstructed suffers from a smear of structure in the line of
sight direction even in the absence of shape noise, and in contrast, the
adaptive lasso algorithm efficiently removes the line of sight smear.
2) The algorithm is able to detect halo with minimal mass limits of $10^{14.0}
M_{\odot}/h$, $10^{14.7} M_{\odot}/h$, $10^{15.0} M_{\odot}/h$ for the low
($z<0.3$), median ($0.3\leq z< 0.6$) and high ($0.6\leq z< 0.85$) redshifts,
respectively, with an average false detection of 0.022/deg$^2$.
3) The estimated redshift of the halos detected from the reconstructed mass
maps are lower than the true redshift by about $0.03$ for halos at low
redshifts ($z\leq 0.4$). The relative redshift bias is below $0.5\%$ for halos
at $0.4<z\leq 0.85$. The standard deviation of the redshift estimation is $0.092$.
\end{abstract}

@@

\begin{figure*}[!t] \includegraphics[width=1.\textwidth]{nfwlet-atom-2D.pdf}
    \caption{The smoothed pixelized basis atoms. The upper row shows the basis
        atoms in Fourier space, and the lower row shows the basis atoms in Real
        space.  The leftmost column is the point mass atom, and the other
        columns are the multi-scale NFW atoms.  The smoothing kernel is
        Gaussian with a standard dev iation of $1.5$ pixels.
    } \label{fig_atoms2D}
\end{figure*}

@@

Weak lensing refers to the phenomenon that light from distant galaxies is
weakly yet coherently distorted by the intervening inhomogeneous density
distribution along the line of sight due to gravity's influence.  As a result,
the shapes of the background galaxies are distorted and the information of the
foreground mass distribution is imprinted to the background galaxy images. Weak
lensing offers a direct probe into the mass distribution, including both
visible matter and invisible dark matter, in our universe \citep[see][for
recent reviews]{revKilbinger15,revRachel17}.
Several large-scale surveys focus on studying the weak-lensing effect at a high
precision level (e.g., HSC \citep{HSC1-data}, KIDS \citep{KIDS13}, DES
\citep{DES05}, LSST \citep{LSSTScienceBook}, Euclid \citep{Euclid2011}, NGRST
\citep{WFIRST15}).

@@

The reconstruction of density map from shear measurements has received
considerable interest since
we can constrain the cosmology model using the peak counts, which are
identified from the weak lensing mass map
\citep{WL-massMap-peakcounts-Jain2000,WL-massMap-peakcountsAna-Fan2010,WL-massMap-peakcountsFM-Lin2016}.
Moreover, we can directly detect massive clusters through
identifying high signal-to-noise (SNR) peaks on the mass map without any
reference to the mass-to-light ratio
\citep{WL-massMap-clusDet-Schneider1996,WL-massMap-clusDet-Hamana2004}.

The two-dimensional ($2$-D) density map reconstruction which recovers an
integration of projected mass along the line of sight has been well studied
within the community
\citep{massMap-KS1993,WL-massMap-Glimpse2D-Lanusse2016,sparseBaysianMassMap-Price2020}
and applied to large-scale surveys
\citep{HSC1-massMaps,massMapDES-Chang2018,DES-SV-massMap-sparsity}.
Cluster detection from $2$-D mass maps has been widely studied and applied to
these large-scale weak-lensing surveys
\citep{WL-massMap-clusDet-CFHT-Shan2012,WL-massMap-clusDet-HSC-Miyazaki2018,WL-massMap-clusDet-HSC-Hamana2020};
however, cross matching with some (e.g., optically selected) cluster catalog is
needed to assign the redshift and extract its physical parameter (such as mass)
for each cluster selected from $2$-D mass map.

On the other hand, we can directly reconstruct $3$-D mass maps taking advantage
of photometric redshift (photo-$z$) information
\citep{LSS-massMap-Wiener-Simon2009,WL-massMap-VanderPlas2011}; however, these
methods either do not have enough spatial resolution to identify clusters, or
they suffer from smears along the line of sight direction, which need to be
overcome for practical searches of clusters in 3D mass maps.  Alternatively,
\citet{WL-clusDet-Hennawi2005} propose to perform a maximum-likelihood
detection of clusters, by convolving tomographic shear measurements with $3$-D
filters that match the tangential shears induced by multi-scale NFW halos,
without fully reconstruct the $3$-D mass maps.

In this paper, we develop a novel method for high-resolution $3$-D mass map
reconstruction that are free from line of sight smears.  The $3$-D
mass map is modeled as a summation of the NFW \citep{halo-NFW1997ApJ} basis
atoms, which have $2$-D multi-scale NFW surface density profiles on the
transverse plane and one-dimensional Dirac delta functions in the line of sight
direction. The adaptive lasso algorithm \citep{AdaLASSO-Zou2006} is applied to
find a sparse reconstruction of the mass maps.

@@

Then, we test our algorithm using the simulations with different setups of the
regularization parameter. We also compare the results of our algorithm, which
uses the NFW dictionary (Section \ref{subsec_test-nfw}), with the point mass
dictionary (Section \ref{subsec_test-pm}).

@@

We have developed a novel method to reconstruct high-resolution $3$-D density
contrast maps from weak-lensing shear measurements and photometric redshift
estimations.  Our method models $3$-D density contrast maps as summations of
NFW atoms with different comoving radii.  With the prior assumption that the
clumpy masses sparsely distribute in the $3$-D space, the density map is
reconstructed using the adaptive lasso algorithm \citep{AdaLASSO-Zou2006}.
We find that the lasso algorithm's solution suffers from a smear of structure
in the line of sight direction even in the absence of galaxy shape noise, and
the adaptive lasso algorithm efficiently removes the line of sight smear of
structure.

We study the performance of cluster detection from the reconstructed $3$-D mass
maps using simulations that apply shear distortions from isolated NFW halos to
HSC-like shapes and photo-$z$ uncertainties.  We conclude that the algorithm is
able to detect halo with minimal mass limits of $10^{14.0} M_{\odot}/h$,
$10^{14.7} M_{\odot}/h$, $10^{15.0} M_{\odot}/h$ for the low ($z<0.3$), median
($0.3\leq z< 0.6$) and high ($0.6\leq z< 0.85$) redshifts, respectively, with an
average false detection of 0.022 deg$^{-2}$. The estimated redshifts of the halos
detected from the reconstructed mass maps are lower than the true redshift by
about $0.03$ for halos at low redshifts ($z\leq 0.4$).  The relative redshift
bias is below $0.5\%$ for halos at $0.4<z\leq 0.85$. The standard deviation of
the redshift estimation is $0.092$.

@@

\section*{Acknowledgements}
XL thank Yin Li and Jiaxin Han for useful discussions.

XL was supported by Global Science Graduate Course (GSGC) program of
University of Tokyo and JSPS KAKENHI (JP19J22222).
\bibliographystyle{aasjournal}
\bibliography{citation}

>>>>

\begin{document}
\title{A new method cluster Detection from three-dimensional weak lensing mass map}
% \author{Xiangchong Li}
% \affiliation{Department of Physics, University of Tokyo, Tokyo 113-0033, Japan}
% \affiliation{Kavli Institute for the Physics and Mathematics of the Universe (WPI),\\
% University of Tokyo, Kashiwa 277-8583, Japan}
% \author{Masamune Oguri}
% \affiliation{Department of Physics, University of Tokyo, Tokyo 113-0033, Japan}
% \affiliation{Kavli Institute for the Physics and Mathematics of the Universe
% (WPI),\\
% University of Tokyo, Kashiwa 277-8583, Japan}
% \affiliation{Research Center for the Early Universe, University of Tokyo, Tokyo
% 113-0033, Japan}
% \author{Wentao Luo}
% \affiliation{Kavli Institute for the Physics and Mathematics of the Universe
% (WPI),\\
% University of Tokyo, Kashiwa 277-8583, Japan}
% \email{xiangchong.li@ipmu.jp}

\begin{abstract}
A new method is developed to reconstruct high-resoluion three-dimensional mass
maps from photometric weak-lensing shear measurements. The three-dimensional
mass map is modeled as a summation of the NFW basis atoms, which have
two-dimensional multi-scale NFW surface density profiles on the transverse
plane and one-dimensional Dirac delta functions in the line of sight direction.
The adaptive lasso algorithm is applied to find a sparse reconstruction of the
mass maps.
We study the performance of three-dimensional detection of clusters of galaxies
from the reconstructed mass map with simulations that apply shear distortions
from isolated halo to HSC-like galaxy shapes with realistic photometric
redshift uncertainties .
We find that the adaptive lasso algorithm efficiently removes the line of sight
smearing, in contrast to the lasso reconstruction that suffers from  the line
of sight smearing even in the absence of shape noise.
The algorithm is able to detect halo with minimal mass limits of $10^{14.0}
h^{-1}M_{\odot}$, $10^{14.7} h^{-1}M_{\odot}$, $10^{15.0} h^{-1}M_{\odot}$ for the low
($z<0.3$), median ($0.3\leq z< 0.6$) and high ($0.6\leq z< 0.85$) redshifts,
respectively, with an average false detection of 0.022/deg$^2$.
The estimated redshift of the halos detected from the reconstructed mass maps
are slightly lower than the true redshift by about $0.03$ for halos at low
redshifts ($z\leq 0.4$). The relative redshift bias is below $0.5\%$ for halos
at $0.4<z\leq 0.85$. The standard deviation of the redshift estimation is
$0.092$.
\end{abstract}

@@

\begin{figure*}[!t] \includegraphics[width=1.\textwidth]{nfwlet-atom-2D.pdf}
    \caption{The smoothed pixelized basis atoms. The upper row shows the basis
        atoms in Fourier space, and the lower row shows the basis atoms in Real
        space.  The leftmost column is the point mass atom, and the other
        columns are the multi-scale NFW atoms.  The smoothing kernel is
        Gaussian with a standard deviation of $1.5$ pixels.
    } \label{fig_atoms2D}
\end{figure*}

@@

Weak lensing refers to the phenomenon that light from distant galaxies is
weakly yet coherently distorted by the intervening inhomogeneous density
distribution along the line of sight due to the influence of gravity.  As a
result, the shapes of the background galaxies are distorted such that the
information of the foreground mass distribution is imprinted to the background
galaxy images. Weak lensing offers a direct probe into the mass distribution,
including both visible matter and invisible dark matter, in our universe
\citep[see][for recent reviews]{revKilbinger15,revRachel17}.
Several large-scale surveys focus on studying the weak-lensing effect at a high
precision level e.g., HSC \citep{HSC1-data}, KIDS \citep{KIDS13}, DES
\citep{DES05}, LSST \citep{LSSTScienceBook}, Euclid \citep{Euclid2011}, NGRST
\citep{WFIRST15}.

@@

The reconstruction of density map from shear measurements has received
considerable interest since
we can constrain the cosmology model using the peak counts, which are
identified from the weak lensing mass map
\citep{WL-massMap-peakcounts-Jain2000,WL-massMap-peakcountsAna-Fan2010,WL-massMap-peakcountsFM-Lin2016}.
Moreover, we can directly detect massive clusters of galaxies through
identifying high signal-to-noise (SNR) peaks on the mass map without any
reference to the mass-to-light ratio
\citep{WL-massMap-clusDet-Schneider1996,WL-massMap-clusDet-Hamana2004}.

The two-dimensional ($2$-D) density map reconstruction which recovers an
integration of projected mass along the line of sight has been well studied
within the community
\citep{massMap-KS1993,WL-massMap-Glimpse2D-Lanusse2016,sparseBaysianMassMap-Price2020}
and applied to large-scale surveys
\citep{HSC1-massMaps,massMapDES-Chang2018,DES-SV-massMap-sparsity}.
Cluster detection from $2$-D mass maps has been widely studied and applied to
these large-scale weak-lensing surveys
\citep{WL-massMap-clusDet-CFHT-Shan2012,WL-massMap-clusDet-HSC-Miyazaki2018,WL-massMap-clusDet-HSC-Hamana2020};
however, cross matching with some (e.g., optically selected) cluster catalog is
needed to assign the redshift and extract its physical parameter (such as mass)
for each cluster selected from $2$-D mass maps.

On the other hand, we can directly reconstruct three-dimensional ($3$-D) mass
maps taking advantage of photometric redshift (photo-$z$) information
\citep{LSS-massMap-Wiener-Simon2009,WL-massMap-VanderPlas2011}; however, these
methods either do not have enough spatial resolution to identify clusters, or
they suffer from smearing along the line of sight direction, which need to be
overcome for practical searches of clusters in 3D mass maps.  Alternatively,
\citet{WL-clusDet-Hennawi2005} propose to perform a maximum-likelihood
detection of clusters, by convolving tomographic shear measurements with $3$-D
filters that match the tangential shears induced by multi-scale NFW halos,
without fully reconstruct the $3$-D mass maps.

In this paper, we develop a novel method for high-resolution $3$-D mass map
reconstruction that are free from line of sight smearing.  The $3$-D
mass map is modeled as a summation of the NFW \citep{halo-NFW1997ApJ} basis
atoms, which have $2$-D multi-scale NFW surface density profiles on the
transverse plane and one-dimensional Dirac delta functions in the line of sight
direction. The adaptive lasso algorithm \citep{AdaLASSO-Zou2006} is applied to
find a sparse reconstruction of the mass maps.

@@

We then test our algorithm using the simulations with different setups of the
regularization parameter. We also compare the results of our algorithm, which
uses the NFW dictionary (Section \ref{subsec_test-nfw}), with the point mass
dictionary (Section \ref{subsec_test-pm}).

@@

We have developed a novel method to reconstruct high-resolution $3$-D density
contrast maps from weak-lensing shear measurements and photometric redshift
estimations.  Our method models $3$-D density contrast maps as summations of
NFW atoms with different comoving radii.  With the prior assumption that the
clumpy masses sparsely distribute in the $3$-D space, the density map is
reconstructed using the adaptive lasso algorithm \citep{AdaLASSO-Zou2006}.
We find that the lasso algorithm uffers from smearing of structure
in the line of sight direction even in the absence of galaxy shape noise, and
the adaptive lasso algorithm efficiently removes the line of sight smearing of
structure.

We study the performance of cluster detection from the reconstructed $3$-D mass
maps using simulations that apply shear distortions from isolated NFW halos to
HSC-like shapes and photo-$z$ uncertainties.  We conclude that the algorithm is
able to detect halo with minimal mass limits of $10^{14.0} h^{-1}M_{\odot}$,
$10^{14.7} h^{-1}M_{\odot}$, $10^{15.0} h^{-1}M_{\odot}$ for the low ($z<0.3$),
median ($0.3\leq z< 0.6$) and high ($0.6\leq z< 0.85$) redshifts, respectively,
with an average false detection of 0.022 deg$^{-2}$. The estimated redshifts of
the halos detected from the reconstructed mass maps are slightly lower than the
true redshift by about $0.03$ for halos at low redshifts ($z\leq 0.4$).  The
relative redshift bias is below $0.5\%$ for halos at $0.4<z\leq 0.85$. The
standard deviation of the redshift estimation is $0.092$.

@@

\section*{Acknowledgements}
We thank Yin Li and Jiaxin Han for useful discussions.

XL was supported by Global Science Graduate Course (GSGC) program of University
of Tokyo and JSPS KAKENHI (JP19J22222).

This work was supported in part by Japan Science and Technology Agency (JST)
CREST JPMHCR1414, and by JST AIP Acceleration Research Grant No. JP20317829,
Japan. This work was supported in part by the World Premier International
Research Center Initiative (WPI Initiative), MEXT, Japan, and JSPS KAKENHI
Grant Nos. JP18K03693, JP20H00181, JP20H05856.

\bibliographystyle{aasjournal}
\bibliography{citation}
